---
title: "Explore and Summarize Data"
author: "Mamadou Diallo"
date: "28 december 2017"
output: html_document
---

### Questions and Issues
[ ] handling outliers

[x] simplify model -> Quality: above 6
[ ] Measure quality of the model: AUC -> NO
[x] training model vs testing model: 70% - 30%
[ ] Use of cross validation
example; https://www.kaggle.com/ash316/eda-to-prediction-dietanic
[ ] use of confusion matrix or AUC
Given the class imbalance ratio, one of the recommend measures for model evaluation is the Area Under the
Precision-Recall Curve (AUPRC), since Confusion matrix accuracy is not meaningful for unbalanced
classification
[ ] handle unbalanced data/skewed classes 
Precision-Recall is a useful measure of success of prediction when the classes are very imbalanced
[ ] variable importance?
[x] correlation among features?
[ ] type of features?
[ ] create function for repetitive tasks: e.g. sampling of data

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo = FALSE, message = FALSE, warning = FALSE}
library(dplyr) # data handling
library(ggplot2) # visualization
library(GGally) # TBD
library(scales) # TBD
library(memisc) # TBD
library(reshape2) # TBD
library(RColorBrewer) # TBD
library(corrplot) # heatmap
library(caTools) # Sampling 
library(rpart) # Decision tree
library(rpart.plot) # plot decision tree
library(caret)
library(e1071)
library(Hmisc) # Histograms
library(gridExtra) # plot multiple plots in one graph
library(ROCR) # ROC curves
```

#### Dataset:
This tidy data set contains 1,599 red wines with 11 variables on the chemical properties of the wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).

#### Guiding question:
Which chemical properties influence the quality of red wines?

```{r Load_the_Data, echo = FALSE}
# Load the data
red <- read.csv("wineQualityReds.csv")
# store it in data frame
df <- tbl_df(red)
# List all column names
names(df)
```
```{r data summary,echo = FALSE}
# Analyse data
str(df)
# Data Summary
summary(df)
```

```{r simplify model into univariate regression model, echo = FALSE}
# no use of feature X: it is an index
df <- subset(df, select = -X)
# simplify model
df$good<-ifelse(df$quality > 6, 1, 0)
```

```{r split into train and test data,echo = FALSE}
set.seed(1011)
msk <- sample.split(df$good, SplitRatio = 0.7)
train <- subset(df, msk == T)
test <- subset(df, msk == F)
```

```{r train table, echo = FALSE}
table(train$good)
# size of the training dataset
dim(train)
```

```{r test table, echo = FALSE}
table(test$good)
# size of test dataset
dim(test)
```

```{r check ratio, echo = FALSE}
14 / (14 + 480)
209 / (209 + 1311)
```


### deal with outliers
```{r}
#df = subset(df,free.sulfur.dioxide<40)
#df = subset(df,volatile.acidity<1.0)
#df = subset(df,fixed.acidity<14.0)
```


```{r Heatmap, echo = FALSE}
#Correlation Heatmap of Variables
corrplot(cor(df),title = "Heatmap")
```

### Interpretation

good and alcohol have a linear relationship (increasing). It is Interesting we find correlation between the dependant variable and an independant variable. It is a good thing since we're trying to predict the independant variable from the independant variables

fixed acidity and citric acidity have a linear relationship (increasing). 

fixed acidity and pH have a linear relationship (decreasing). 

fixed acidity and density have a linear relationship (increasing).

volatile acidity and citric acidity have a linear relationship (decreasing).

free.sulfur.dioxide and total.sulfure.dioxide have a linear relationship.

Since it is not suffucient to look at the correlation numbers, we need to look at the graphs

### Interpretation
TBD
```{r Histograms of each continuous features for the global dataset, echo = FALSE}
hist.data.frame(df[, 1:11],mtitl = "Histograms of each continuous features")
```



```{r Discretizing continuous features, echo = FALSE}
discreteContFeature <- function(featureTable,featureName, start = 0.01) {
  # compute quantiles
  q <- quantile(featureTable,seq(0, 1, by = 0.1))
  # the 2 wines belong to the first interval: make them part of the first interval  
  q[1] = q[1] - start
  #print(table(featureTable))
  # create intervals open to the left and closed to the right
  qvar <- cut(featureTable, q)
  tab <- table(qvar, df$good)
  # ploting the bar plot
  t1 = prop.table(tab, 1)[, 2]
  names(t1) = c("NOK", "OK")
  #  p = barplot(t(t1),ylim=c(0,0.7),las=3,main=featureName,ylab="rate of good wine",density = 0)
  p = barplot(t(t1),las = 3, main = featureName, ylab = "rate of good wine", density = 0)
  # plot average wine quality
  abline(h = 0.1357, lty = 2)

  #return(prop.table(tab,1))
  return("Plot done")
} 
```

```{r Discretize features, echo = FALSE}
discreteContFeature(df$alcohol, "Alcool", 0.01)
discreteContFeature(df$volatile.acidity, "Volatile Acidity", 0.01)
discreteContFeature(df$fixed.acidity, "Fixed Acidity", 0.01)
discreteContFeature(df$total.sulfur.dioxide, "Total Sulfur Dioxide", 0.01)
discreteContFeature(df$density, "Density", 0.01)
discreteContFeature(df$sulphates, "Sulphates", 0.01)
discreteContFeature(df$citric.acid, "citric acid", 0.01)
discreteContFeature(df$chlorides,"chlorides", 0.01)
```


```{r Histogram Quality feature, echo = FALSE}
table(df$quality)
ggplot(data = df,aes(x = quality)) +
  geom_histogram(stat = "count", color = 'black', fill = 'orange') +
  ggtitle("Histogram Quality feature")
```
```{r Histogram good feature, echo = FALSE}
table(df$good)
ggplot(data = df,aes(x = good)) +
  geom_histogram(stat = "count", color = 'black', fill = 'orange') +
  ggtitle("Histogram good feature")
```


```{r echo = FALSE}
by(df$volatile.acidity,df$good, summary)
```

> **Tip**: The dependant variable is "quality" as an ordinal variable
Continuous features: All but "quality"
This is a classification problem.

Issue: the data is not balanced in wine quality.

```{r histogram of volatile acidity for training dataset, echo = FALSE}
ggplot(data = train,aes(x = volatile.acidity)) +
  geom_histogram(colour = 'black', fill = 'orange', bins = 30) +
  scale_x_continuous() +
  facet_wrap(~good,ncol = 1, scales = "free_y") +
  ggtitle("histogram of volatile acidity")
```

```{r histogram of volatile acidity, echo = FALSE}
ggplot(data = df,aes(x = volatile.acidity)) +
  geom_histogram(bins = 30, color = 'black', fill = 'orange') +
  scale_x_continuous() +
  ggtitle("histogram of volatile acidity")
```
```{r histogram of volatile acidity log-normal, echo = FALSE}
ggplot(data = df,aes(x = volatile.acidity)) +
  geom_histogram(bins = 30, color = 'black', fill = 'orange') +
  scale_x_log10() +
  ggtitle("histogram of volatile acidity log-normal")
```


```{r histogram of citric acid, echo = FALSE}
ggplot(data = df,aes(x = citric.acid)) +
  geom_histogram(bins = 50, color = 'black', fill = 'orange') +
  scale_x_continuous() +
  facet_wrap(~good,ncol = 1,scales = "free_y") + 
  ggtitle("histogram of citric acid")
```
```{r Histogram of residual sugar, echo = FALSE}
ggplot(data = df,aes(x = residual.sugar)) +
  geom_histogram( bins = 30, color = 'black', fill = 'orange') +
  scale_x_continuous() +
  ggtitle("Histogram of residual sugar")
```
```{r histogram of residual sugar log-normal, echo = FALSE}
ggplot(data = df,aes(x = residual.sugar)) +
  geom_histogram(color = 'black', fill = 'orange') +
  scale_x_log10() + 
  ggtitle("histogram of residual sugar log-normal")
```
```{r histogram of residual sugar, echo = FALSE}
ggplot(data = df,aes(x = residual.sugar)) +
  geom_histogram(color = 'black', fill = 'orange') +
  scale_x_continuous() +
  ggtitle("histogram of residual sugar")
```
```{r echo = FALSE}
ggplot(aes(x = residual.sugar, y = ..count../sum(..count..)),
       data = df) +
  geom_freqpoly() +
  scale_x_log10() +
  xlab('Residual Sugar') +
  ylab('Proportion of wines with that residual sugar count')
```


```{r boxplot of good pH vs good, echo = FALSE}
ggplot(data = df, mapping = aes(x = good, y = pH)) +
  geom_boxplot()
```

```{r Histogram of pH,echo = FALSE}
ggplot(data = df,aes(x = pH))+
  geom_histogram(colour = 'black', fill = 'orange') +
  scale_x_continuous() +
  facet_wrap(~good, ncol = 1) +
  ggtitle("Histogram of pH")
```

```{r boxplot of density vs good, echo = FALSE}
ggplot(data = df,mapping = aes(x = good, y = density)) +
  geom_boxplot() +
  ggtitle("boxplot of density vs good")
```


```{r boxplot of citric acid vs good, echo = FALSE}
ggplot(data = df, mapping = aes(x = good, y = citric.acid)) +
  geom_boxplot() +
  ggtitle("boxplot of citric acid vs good")
```



```{r echo = FALSE, The_Matrix, echo = FALSE}
#ggpairs(df,
#  lower = list(continuous = wrap("points", shape = I('.'))),
#  upper = list(combo = wrap("box", outlier.shape = I('.'))))
```


### Study 2 variables
```{r scatter plot citric acid vs fixed acidity, echo = FALSE}
ggplot(aes(x = fixed.acidity, y = citric.acid), data = df) +
  geom_point(alpha = 1/10,
             position = position_jitter(height = 0),
             color = 'red') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_smooth(method = "lm", color = "blue") +
  ggtitle("scatter plot citric acid vs fixed acidity")
```
#### Intepretation:
the fixed acidity is increasing with the citric acid concentration
Correlation = TBD

```{r scatter plot fixed acidity vs pH, echo = FALSE}
ggplot(aes(x = fixed.acidity, y = pH), data = df) +
  geom_point(alpha = 1/10,
             position = position_jitter(height = 0),
             color = 'red') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_smooth(method = "lm", color = "blue") +
  ggtitle("scatter plot fixed acidity vs pH")
```
#### Interpretation
The fixed acidity is increasing as the pH is decreasing

Correlation = TBD

```{r scatter plot fixed acidity vs density, echo = FALSE}
# Create a scatterplot of fixed acidity vs density
# and omit the top 5% of fixed acidity vs density
# values.
ggplot(aes(x = fixed.acidity, y = density), data = df) +
  geom_point(alpha = 1/10,
             position = position_jitter(height = 0),
             color = 'red') +
  xlim(4.60, quantile(df$fixed.acidity, 0.95)) +
  ylim(0.99, quantile(df$density, 0.95)) +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_smooth(method = "lm",color = "blue") +
  ggtitle("scatter plot fixed acidity vs pH")
```
#### Interpretation:
fixed.acidity is increasing with density
Correlation = TBD


```{r scatter plot free sulfur dioxide vs total sulfur dioxide, echo = FALSE}
# Create a scatterplot of free sulfure dioxide vs total sulfur dioxide
# and omit the top 5% of free sulfure dioxide vs total sulfur dioxide
# values.
ggplot(aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide), data = df) +
  geom_point(alpha = 1/10,
             position = position_jitter(height = 0),
             color = 'red') +
  xlim(1, quantile(df$free.sulfur.dioxide, 0.95)) +
  ylim(6,quantile(df$total.sulfur.dioxide,0.95)) +
  geom_line(stat = 'summary',fun.y=mean) +
  geom_smooth(method = "lm", color = "blue") +
  ggtitle("scatter plot free sulfur dioxide vs total sulfur dioxide")
```
#### Interpretation
free.sulfur.dioxide increases with total.sulfur.dioxide
Correlation = TBD

```{r group by free sulfur dioxide, echo = FALSE}
df.by_free_sulfur <- df %>% group_by(free.sulfur.dioxide) %>%
  summarise(total_sulfur_mean = mean(total.sulfur.dioxide),
            total_sulfur_median = median(total.sulfur.dioxide),
            n = n()) %>%
  arrange(free.sulfur.dioxide)

head(df.by_free_sulfur)
```
```{r}
ggplot(aes(x = free.sulfur.dioxide, y = total_sulfur_mean), data = df.by_free_sulfur) +
  geom_line()
```


```{r scatter plot volatile acidity vs citric acid, echo = FALSE}
ggplot(aes(x = volatile.acidity, y = citric.acid), data = df) +
  geom_point(alpha = 1/10,
             color = 'red') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_smooth(method = "lm",color = "blue") +
  ggtitle("scatter plot volatile acidity vs citric acid")
```
#### Interpretation
the volatile acidity decreases as citric acid increase
Correlation = TBD


```{r}
names(train)
```
```{r keeping most significant features, echo = FALSE}
#train <- subset(train, select = c(pH, volatile.acidity, citric.acid, total.sulfur.dioxide, sulphates, alcohol, good))
#test<- subset( test, select = c(pH, volatile.acidity, citric.acid, total.sulfur.dioxide, sulphates, alcohol))
```


### Decision Tree model
```{r Decision Tree model, echo = FALSE}
tree.model <- rpart(good ~ ., data = train, method = "class")
prp(tree.model) 
```

```{r}
tree.predict <- predict(tree.model, test, type = "class")
#confusionMatrix(test$good, tree.predict)
```

### Logistic Regression
Baseline model accuracy: if we would predict that all wine are not good (most frequent outcome) = 0.86
We'll try to beat this accuracy with the logistical model
```{r}
table(df$good)
1382 / (1382 + 217)
```
```{r Log model}
wineLog1 <- glm(good ~ alcohol, data = train, family = binomial)
```
```{r coefficients}
summary(wineLog1)
```

```{r}
wineLog2 <- glm(good ~ alcohol + volatile.acidity, data = train, family =binomial)
```

```{r}
summary(wineLog2)
```

```{r}
wineLog3 <- glm(good ~ alcohol + volatile.acidity + citric.acid, data = train, family =binomial)
summary(wineLog3)
```

```{r}
wineLog4 <- glm(good ~ alcohol + volatile.acidity + total.sulfur.dioxide + sulphates, data = train, family = binomial)
summary(wineLog4)
```
### Interpretation
correlations !!

```{r Prediction}
predTrain4 <- predict(wineLog4, type = "response")
```
```{r}
summary(predTrain4)
```
```{r average predictions for each of the true outcomes}
tapply(predTrain4, train$good, mean)
```
```{r}
table(train$good, predTrain4 > 0.2)
```
```{r}
ROCRpred <- prediction(predTrain4, train$good)
```
```{r}
ROCRperf <- performance(ROCRpred, "tpr", "fpr")
```
```{r}
plot(ROCRperf)
```
```{r}
plot(ROCRperf, colorize = T)
```
```{r}
names(train)
```

```{r}
m1 <- lm(good ~ alcohol, data = train)
m2 <- update(m1, ~ . + sulphates)
m3 <- update(m2, ~ . + density)
m4 <- update(m3, ~ . + residual.sugar)
m5 <- update(m4, ~ . + volatile.acidity)
m6 <- update(m5, ~ . + chlorides)
m7 <- update(m6, ~ . + fixed.acidity)
m8 <- update(m7, ~ . + citric.acid)
m9 <- update(m8, ~ . + free.sulfur.dioxide)
m10 <- update(m9, ~ . + pH)
m11 <- update(m10, ~ . + total.sulfur.dioxide)
mtable(m1, m2, m3, m4, m5, m6, m7, m8, m9, m10, m11)
```
## Interpretation
density coefficient flipped from positive to negative 

```{r}
ggplot(aes(x = citric.acid, y = fixed.acidity,color = factor(good)), data = df) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE) +
  ggtitle("scatter plot citric acid vs fixed acidity")
```


### Predictions
2)Support Vector Machines
3)Random Forest
4)K-Nearest Neighbours
5)Naive Bayes
[x] Decision Tree


### Resources:

Data set location:
https://www.google.com/url?q=https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityReds.csv&sa=D&ust=1515532515282000&usg=AFQjCNHpiAAlB8pRni-QOCbCQLSZ5yYCkw

coding in R
R Graphics Cookbook - Winston Chang

https://www.kaggle.com/grosvenpaul/beginners-guide-to-eda-and-random-forest-using-r
https://www.kaggle.com/meepbobeep/intro-to-regression-and-classification-in-r
https://www.youtube.com/watch?v=jlOj2THP4xc
geography matters
yeast
weather matters
the smell matters
Age?
Movie: Somm (2012) on Netflix -> done
Source for handling unbalanced data/skewed classes:
https://www.datasciencecentral.com/profiles/blogs/data-science-with-python-exploratory-analysis-with-movie-ratings?utm_content=buffer5a99e&utm_medium=social&utm_source=twitter.com&utm_campaign=buffer

Edx Course: The Analytics Edge - Course "The Statistical Sommelier"
https://d37djvu3ytnwxt.cloudfront.net/assets/courseware/v1/7347c7fce82a329565388cd40adce2c9/asset-v1:MITx+15.071x_3+1T2016+type@asset+block/Unit2_WineRegression_AllSlides.pdf

Linear regression: Question answered: Can analytics be used to come up with a different system for judging wine? 
Independant variables: Age, Weather - Average growing season Temperature, Harvest rain, Winter rain. R2 = 0.83. most Significant coefficient: AGS, Harvest rain 
Techniques used for unbalanced data:
Oversampling: 
Under-sampling:

Repo1:
https://github.com/GeorgeVince/DAND_White-Wine-EDA/blob/master/white_wine_analysis.html


REFLECTIONS:
struggles, 
unbalanced data
build model with minimum features
validation of the model
multicolinearity in the model - two independant variables significantly correlated. We have to remove one variable one at a time.

successes,
factors influencing the quality of wine

and ideas for future exploration
take into account of independant features such as: Age, Weather
other prediction models

