---
title: "EDA"
author: "Mamadou Diallo"
date: "28 décembre 2017"
output: html_document
---
#### Variables
   1 - fixed acidity (tartaric acid - g / dm^3)
   2 - volatile acidity (acetic acid - g / dm^3)
   3 - citric acid (g / dm^3)
   4 - residual sugar (g / dm^3)
   5 - chlorides (sodium chloride - g / dm^3
   6 - free sulfur dioxide (mg / dm^3)
   7 - total sulfur dioxide (mg / dm^3)
   8 - density (g / cm^3)
   9 - pH
   10 - sulphates (potassium sulphate - g / dm3)
   11 - alcohol (% by volume)
   Output variable (based on sensory data): 
   12 - quality (score between 0 and 10)

### Questions and Issues
[ ] handling outliers

[x] simplify model -> Quality: above 6
[ ] Measure quality of the model: AUC
[ ] training model vs testing model: 70% - 30%
[ ] Use of cross validation
example; https://www.kaggle.com/ash316/eda-to-prediction-dietanic
[ ] use of confusion matrix or AUC
Given the class imbalance ratio, one of the recommend measures for model evaluation is the Area Under the
Precision-Recall Curve (AUPRC), since Confusion matrix accuracy is not meaningful for unbalanced
classification
[ ] handle unbalanced data
techniques see end of the document
[ ] variable importance?
[x] correlation among features?
[ ] type of features?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages,echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr) # data handling
library(ggplot2) # visualization
library(GGally) # TBD
library(scales) # TBD
library(memisc) # TBD
library(reshape2) # TBD
library(RColorBrewer) # TBD
library(corrplot) # TBD
library(caTools)
```


```{r Load_the_Data,echo=FALSE}
# Load the data
red <- read.csv("wineQualityReds.csv")
# store it in data frame
df <- tbl_df(red)
# List all column names
names(df)
```
```{r split into train and test data}
set.seed(1011)
msk <- sample.split(df,SplitRatio = 0.7)
train <- subset(df,msk == T)
test <- subset(df,msk == F)
table(train$quality)
table(test$quality)
dim(train)
dim(test)
```
```{r}
492/1107
```

```{r data_wrangling,echo=FALSE}
# no use of feature X
df <- subset( df, select = -X )
# simplify model
df$good<-ifelse(df$quality>6,1,0)
# Analyse data
str(df)
# Data Summary
summary(df)
```

```{r Heatmap}
#Correlation Heatmap of Variables
corrplot(cor(df))
```

```{r change quality into factor,echo=FALSE}
df$quality = factor(df$quality)
df$good = factor(df$good)
```
### Interpretation
TBD

```{r Histogram Quality,echo=FALSE}
table(df$quality)
ggplot(data = df,aes(x = quality))+
  geom_histogram(stat = "count")
```
```{r}
table(df$good)
ggplot(data = df,aes(x = good))+
  geom_histogram(stat = "count")
```


```{r}
by(df$fixed.acidity,df$quality,summary)
```
```{r}
by(df$volatile.acidity,df$good,summary)
```

> **Tip**: The dependant variable is "quality" as an ordinal variable
Continuous features: All but "quality"
This is a classification problem.

Issue: the data is not balanced in wine quality.


```{r Histogram fixed.acidity,echo=FALSE}
ggplot(data = df,aes(x = fixed.acidity))+
  geom_histogram(colour='white',fill='black')+
  scale_x_continuous()+
  facet_wrap(~quality)
```
```{r}
ggplot(data = df,aes(x = volatile.acidity))+
  geom_histogram(colour='white',fill='black')+
  scale_x_continuous()+
  facet_wrap(~good)
```

```{r}
ggplot(data = df,aes(x = volatile.acidity))+
  geom_histogram(bins=30,fill='black',colour='white')+
  scale_x_continuous()
```
```{r}
ggplot(data = df,aes(x = volatile.acidity))+
  geom_histogram(bins=30,fill='black',colour='white')+
  scale_x_log10()
```


```{r}
ggplot(data = df,aes(x = citric.acid))+
  geom_histogram(bins=50,fill='black',colour='white')+
  scale_x_continuous()
```
```{r residual.sugar}
ggplot(data = df,aes(x = residual.sugar))+
  geom_histogram(bins=16,fill='black',colour='white')+
  scale_x_continuous()
```
```{r}
ggplot(data = df,aes(x = residual.sugar))+
  geom_histogram(fill='black',colour='white')+
  scale_x_log10()
```
```{r chlorides}
ggplot(data = df,aes(x = residual.sugar))+
  geom_histogram(fill='black',colour='white')+
  scale_x_continuous()
```

```{r echo=FALSE}
ggplot(data = df,mapping = aes(x = quality,y = pH))+
  geom_boxplot()
```
```{r}
ggplot(data = df,mapping = aes(x = good,y = pH))+
  geom_boxplot()
```

```{r Histogram fixed.acidity,echo=FALSE}
ggplot(data = df,aes(x = pH))+
  geom_histogram(colour='white',fill='black')+
  scale_x_continuous()+
  facet_wrap(~quality)
```

```{r echo=FALSE}
ggplot(data = df,mapping = aes(x = quality,y = alcohol))+
  geom_boxplot()
```
```{r echo=FALSE}
ggplot(data = df,mapping = aes(x = quality,y = density))+
  geom_boxplot()
```
```{r}
ggplot(data = df,mapping = aes(x = good,y = density))+
  geom_boxplot()
```

```{r echo=FALSE}
ggplot(data = df,mapping = aes(x = quality,y = citric.acid))+
  geom_boxplot()
```
```{r}
ggplot(data = df,mapping = aes(x = good,y = citric.acid))+
  geom_boxplot()
```

```{r echo=FALSE}
ggplot(data = df,mapping = aes(x = quality,y = volatile.acidity))+
  geom_boxplot()
```

```{r echo=FALSE, The_Matrix}
ggpairs(df,
  lower = list(continuous = wrap("points", shape = I('.'))),
  upper = list(combo = wrap("box", outlier.shape = I('.'))))
```



```{r echo=FALSE}
ggplot(aes(x = pH, y = quality,colour=alcohol), data = df) + 
  geom_point(alpha = 0.5, size = 1) +
  ggtitle('Quality vs PH')
```
### Predictions
1)Logistic Regression
2)Support Vector Machines(Linear and radial)
3)Random Forest
4)K-Nearest Neighbours
5)Naive Bayes
6)Decision Tree


### Resources:
https://www.kaggle.com/grosvenpaul/beginners-guide-to-eda-and-random-forest-using-r
https://www.kaggle.com/meepbobeep/intro-to-regression-and-classification-in-r
https://www.youtube.com/watch?v=jlOj2THP4xc
geography matters
yeast
weather matters
the smell matters
Age?
Movie: Somm (2012) on Netflix -> done
Source for handling unbalanced data:
https://www.datasciencecentral.com/profiles/blogs/data-science-with-python-exploratory-analysis-with-movie-ratings?utm_content=buffer5a99e&utm_medium=social&utm_source=twitter.com&utm_campaign=buffer
Techniques used for unbalanced data:
Oversampling: 
Under-sampling:
